Notes on Control Plane Inference

### 2021-03-31 inference.tex
outline p4v. think it doesnt work, because we need to know which row each packet hits in each table

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\pv}{\mathit{p4v}}

@@

@@

\section{P4v comparison}

Ive said before that the P4v-style instrumentation is an overapproximation of
the control plane behavior. It's a slightly inaccurate claim in that it
correctly captures all possible control plane behaviors, but it doesn't have a
handle on constrain control plane behavior.

\begin{figure}
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\pv : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{\pv : \Prog \to \Instr} \\
  \pv(h.f := e) && h.f := e \\
  \pv(\assume b) && \assume b \\
  \pv(\assert b) && \assert b \\
  \pv(c_1;c_2) && \pv(c_1;c_2) \\
  \pv(c_1 \choice c_2) && \pv(c_1) \choice \pv(c_2) \\
  \pv(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
  \pv(t.\apply()) && \\
  \multicolumn{3}{>{\qquad}r}{
    \displaystyle
    \bigchoice_{i = 0}^{|t.\action|-1}t.\action[i]()
  }
  \end{array}\]
  \caption{A \texttt{p4v}-like instrumentation. For simplicity of reasoning
    about variables, assume that actions take no action data. }
  \label{fig:p4vlike}
\end{figure}

\begin{theorem}[\texttt{p4v} over-approximates]
  \label{thm:p4v-over}
  For every $c \in \Prog$, well-formed state $\sigma : \Value \to \Row^* $,
  model $\mu \in \Model$, and packet $\pkt \in \Pkt$,
  $\denote{c}^\sigma\,\mu\,\pkt \subseteq \denote{\pv(c)}^\sigma\,\mu\,\pkt$
\end{theorem}

Whereas
\[\begin{array}{l}
\denote{\pv(t.\apply())}^\emptyset\,\emptyset\,\pkt \\
= \denote{
    \displaystyle
    \bigchoice_{i = 0}^{|t.\action|-1} t.\action[i]()
}^\emptyset\,\emptyset\,\pkt \\
= \denote{\SKIP}^\emptyset\,\emptyset\,\pkt \cup \denote{h.f := 99}^\emptyset\,\emptyset\,\pkt\\
= \{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}
\end{array}\]

>>>>

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\nondet}{\mathsf{nondet}}
\newcommand{\ghostly}{\mathsf{ghostly}}

@@

\section{Ghost State}

Following p4v, we want construct a ghostly header $\gamma$ from a program $p$
and use it to instrument the program to capture all dataplane behaviors.

\[
\begin{array}{l>{\triangleq}cl}
  \multicolumn{3}{l}{\ghostly :: \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{\ghostly :: \Prog \to \GCL} \\
  \ghostly(h.f := e)
  && h.f := e \\
  \ghostly(\assert b)
  && \assert b\\
  \ghostly(\assume b)
  && \assume b \\
  \ghostly(c_1;c_2)
  && \ghostly(c_1);\ghostly(c_2) \\
  \ghostly(c_1\choice c_2)
  && \ghostly(c_1) \choice \ghostly(c_1)\\
  \ghostly(t.\apply())
  &&\\
  \multicolumn{3}{c}{
    \begin{array}[t]{>{\displaystyle\qquad}l}
       \gamma.t_{\mathit{reach}} := [1]_1; \\
       \gamma.t_{\mathit{read}(0)} := t.\keys[0]; \\
       \vdots \\
       \gamma.t_{\mathit{read}(|t.\keys|-1)} := t.\keys[|t.\keys|-1]; \\[1em]
       \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
       \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}{
         \left(\begin{array}{l}
           \gamma.t_{\mathit{hit}} := [1]_1 ;\\
           \gamma.t_{\mathit{action}} := [1]_1;\\
           t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
         \end{array}\right)
       }\right)
     \end{array}}
\end{array}
\]

And now what we want to prove is

\begin{theorem}{Instrumentation sufficiency}
  For every well-formed program $c \in \Cmd$, every state $\sigma : \Value \to
  \Row^*$, there is some $b_\sigma \in \BExpr$ such that for every model $\mu
  \in \Model$ and every $\pkt \in \Pkt$,
  \[\denote{c}^\sigma\,\mu\,\pkt =_{\not\gamma} \denote{\ghostly(c);\assume b}^\sigma\,\mu\,\pkt\]
\end{theorem}

\begin{proof}
  Proceed by induction on $c$:
  \begin{enumerate}[align=left]
  \item[$(h.f:=e)$] Let $b_\sigma = \TRUE$; immediate.
  \item[$(\assume b)$] Let $b_\sigma = \TRUE$; immediate.
  \item[$(\assert b)$] Let $b_\sigma = \TRUE$; immediate.
  \item[$(c_1;c_2)$] By IHs. $b_\sigma = b_1 \wedge b_2$ because $b_1$ and $b_2$ refer to disjoint variables.
  \item[$(c_1 \choice c_2)$] \textit{sim}.
  \item[$(t.\apply())$]

    Let $b_\sigma$ be defined as follows:
    \[
    b_\sigma = \bigvee_{r \in \sigma(t.\id)}
    \begin{array}{>{\displaystyle}l}
      \gamma
    \end{array}
    \]

    We can compute as follows:
    \[\begin{array}{>{\displaystyle}ll}
    \denote{\ghostly(c)}^\sigma\,\mu\,\pkt\\
    =\denote{
        \begin{array}{>{\displaystyle}l}
          \gamma.t_{\mathit{reach}} := [1]_1; \\
          \gamma.t_{\mathit{read}(0)} := t.\keys[0]; \\
          \vdots \\
          \gamma.t_{\mathit{read}(|t.\keys|-1)} := t.\keys[|t.\keys|-1]; \\[1em]
          \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
          \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}
            \left(\begin{array}{l}
              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
              \gamma.t_{\mathit{action}} := j;\\
              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
            \end{array}\right)
          \right)
    \end{array}}^\sigma\,\mu\,\pkt \\[2em]
    =\denote{
        \begin{array}{>{\displaystyle}l}
          \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
          \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}
            \left(\begin{array}{l}
              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
              \gamma.t_{\mathit{action}} := j;\\
              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
            \end{array}\right)
          \right)
    \end{array}}^\sigma\!\mu\,\pkt' \\
    \qquad\qquad \text{where}~\pkt' = \pkt\left\{\begin{array}{l}
    \gamma.t_{\mathit{reach}} \mapsto [1]_1,\\
    \overrightarrow{\gamma.t_{\textit{key}} \mapsto \edenote{t.\keys}\,\pkt\,\mu}
    \end{array}\right\} \\
    =\denote{
      \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)}^\sigma\!\mu\,\pkt'
    \cup
    \denote{\left(\bigchoice_{j = 1}^{|t.\action|-1}
            \left(\begin{array}{l}
              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
              \gamma.t_{\mathit{action}} := j;\\
              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
            \end{array}\right)
            \right)
    }^\sigma\!\mu\,\pkt' \\
    = \left\{(\mu,\pkt'\left\{\gamma.t_{\mathit{hit}} \mapsto [0]_1\right\})\right\}
    \cup
        \denote{\left(\bigchoice_{j = 1}^{|t.\action|-1}
          \left(\begin{array}{l}
            \gamma.t_{\mathit{hit}} := [1]_1 ;\\
            \gamma.t_{\mathit{action}} := j;\\
            t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
          \end{array}\right)
          \right)
        }^\sigma\!\mu\,\pkt' \\
  = \left\{(\mu,\pkt'\left\{\gamma.t_{\mathit{hit}} \mapsto [0]_1\right\})\right\}
  \cup
  \left(\bigcup_{j = 1}^{|t.\action|-1}
  \denote{t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})}^\sigma\,\mu\,\pkt'\left\{
  \begin{array}{l}
    \gamma.t_{\textit{hit}} \mapsto [1]_1,\\
    \gamma.t_{\textit{action}} \mapsto j
  \end{array}
  \right\}
  \right) \\

    \end{array}
\]

\section{Tables as nondeterministic choice}

Ive said before that the nondeterministic choice is an overapproximation of
the control plane behavior. It's a slightly inaccurate claim in that it
correctly captures all possible control plane behaviors, but it doesn't have a
handle on constrain control plane behavior.

\begin{figure}
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\nondet : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{\nondet : \Prog \to \Instr} \\
  \nondet(h.f := e) && h.f := e \\
  \nondet(\assume b) && \assume b \\
  \nondet(\assert b) && \assert b \\
  \nondet(c_1;c_2) && \nondet(c_1;c_2) \\
  \nondet(c_1 \choice c_2) && \nondet(c_1) \choice \nondet(c_2) \\
  \nondet(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
  \nondet(t.\apply()) && \\
  \multicolumn{3}{>{\qquad}r}{
    \displaystyle
    \bigchoice_{i = 0}^{|t.\action|-1}t.\action[i]()
  }
  \end{array}\]
  \caption{A \texttt{p4v}-like instrumentation. For simplicity of reasoning
    about variables, assume that actions take no action data. }
  \label{fig:p4vlike}
\end{figure}

\begin{theorem}[\texttt{p4v} over-approximates]
  \label{thm:p4v-over}
  For every $c \in \Prog$, well-formed state $\sigma : \Value \to \Row^* $,
  model $\mu \in \Model$, and packet $\pkt \in \Pkt$,
  $\denote{c}^\sigma\,\mu\,\pkt \subseteq \denote{\nondet(c)}^\sigma\,\mu\,\pkt$
\end{theorem}

Whereas
\[\begin{array}{l}
\denote{\nondet(t.\apply())}^\emptyset\,\emptyset\,\pkt \\
= \denote{
    \displaystyle
    \bigchoice_{i = 0}^{|t.\action|-1} t.\action[i]()
}^\emptyset\,\emptyset\,\pkt \\
= \denote{\SKIP}^\emptyset\,\emptyset\,\pkt \cup \denote{h.f := 99}^\emptyset\,\emptyset\,\pkt\\
= \{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}
\end{array}\]

@@
